---
title: "Homework 2.1"
author: "Duubar Villalobos Jimenez   mydvtech@gmail.com"
date: "September 16, 2018"
output:
  pdf_document: default
  html_document: default
  prettydoc::html_pretty:
    highlight: github
    theme: leonids
    toc: yes
subtitle: CUNY MSDS DATA 609
---

```{r, echo=FALSE, warning=FALSE, error=FALSE, cache=FALSE, results='hide', message=FALSE}

knitr::opts_chunk$set(echo = TRUE)
type <- 'latex'
```


```{r, echo=FALSE, warning=FALSE, error=FALSE, cache=FALSE, results='hide', message=FALSE}

library(knitr)
library(data.table)
library(xtable)
library(plotly)

```

# Problems

The below problems are taken from the text book:

A First Course in Mathematical Modeling, 5th Edition. Frank R. Giordano, William P. Fox, Steven B. Horton. ISBN-13: 9781285050904.

## Exercise \#2 Page 113.

The following table gives the elongation $e$ in inches per inc(in./in.) for a given stress $S$ on a steel wire measured in pounds per square in (lb/in.$^2$). Test the model $e=c_1S$ by plotting the data. Estimate $c1$ graphically.

```{r, echo=FALSE}
S <- c(5,10,20,30,40,50,60,70,80,90,100)
e <- c(0,19,57,94,134,173,216,256,297,343,390)
t.df <- data.frame(S = S, e = e)
```


```{r, echo=FALSE, results='asis'}
print(xtable(t.df, caption='Elongation $e$ with $S(x10^{-3})$ and $e(x10^5)$.', digits=0), hline.after=c(0,11), include.rownames=FALSE, type=type, comment=FALSE)
```

### Solution

Let's plot our original data and see how it looks.

```{r, echo=FALSE}
plot(t.df$e~t.df$S,
      type="p", 
      col="blue",
      main='Elongation data',
      xlab="S", 
      ylab="e")
```

From the above, we can appreciate that the data is fairly linear. Hence, we could express it in the form $e= c_1S$ as provided above.

Where $c_1 = \frac{390 - 0}{100 - 5} = 4.105263$

thus, it returns the following model $e=4.105263S$

```{r, echo=FALSE}
plot(t.df$e~t.df$S,
        type="p", 
        col="blue",
        main='Elongation data',
        xlab="S", 
        ylab="e")
lines(t.df$S, 4.105263 * t.df$S, col="red", type="l", pch=17)
legend("topleft", 
       c("given data","linear model"),
       fill=c("blue","red")
       )
```

From what we can see, our linear model is not very accurate but it is beyond our current problem.



## Exercise \#2.a Page 121.

For each of the following data sets, formulate the mathematical model that minimizes the largest deviation between the data and the line $y=ax+b$. If a computer is available, solve for the estimates of $a$ and $b$.

```{r, echo=FALSE}
x <- c(1.0,2.3,3.7,4.2,6.1,7.0)
y <- c(3.6,3.0,3.2,5.1,5.3,6.8)
t.df <- data.frame(y,x)
```


```{r, echo=FALSE, results='asis'}
print(xtable(t.df, caption='Problem 2.a data set.', digits=1), hline.after=c(0,6), include.rownames=FALSE, type=type, comment=FALSE)
```

### Solution

Let's have a visual representation of the data set.

```{r, echo=FALSE}
plot(t.df$y~t.df$x,
      type="p", 
      col="blue",
      main='Given data set',
      xlab="x", 
      ylab="y")
```

The idea here is to minimize our largest deviation in our data set and predicted values; hence, we could perform as follows:

**Using Chebyshev criterion**

That is:

$r - (3.6 - (1.0 a + b)) \ge 0$,
$r + (3.6 - (1.0 a + b)) \ge 0$

$r - (3.0 - (2.3 a + b)) \ge 0$,
$r + (3.0 - (2.3 a + b)) \ge 0$

$r - (3.2 - (3.7 a + b)) \ge 0$,
$r + (3.2 - (3.7 a + b)) \ge 0$

$r - (5.1 - (4.2 a + b)) \ge 0$,
$r + (5.1 - (4.2 a + b)) \ge 0$

$r - (5.3 - (6.1 a + b)) \ge 0$,
$r + (5.3 - (6.1 a + b)) \ge 0$

$r - (6.8 - (7.0 a + b)) \ge 0$,
$r + (6.8 - (7.0 a + b)) \ge 0$

The above, is the formulation of the mathematical model that minimizes the largest deviation between the data and the line $y = ax + b$.


**Finding the estimates of $a$ and $b$.**

Using least squares criterion.

```{r}
lm.y <- lm(t.df$y~t.df$x)
summary(lm.y)
```

From th above, we can conclude that our function employing the $R^2$ method is:

$y = 2.2149 + 0.5642x$

```{r, echo=FALSE}
plot(t.df$y~t.df$x,
        type="p", 
        col="blue",
        main='Given data',
        xlab="x", 
        ylab="y")
lines(t.df$x, 2.2149 + 0.5642 * t.df$x,  col="red", type="l", pch=17)
legend("topleft", 
       c("given data","linear model"),
       fill=c("blue","red")
       )
```

# END.