---
title: "Homework 02"
author: "Duubar Villalobos Jimenez   mydvtech@gmail.com"
date: "September 9, 2018"
output:
  pdf_document: default
  html_document: default
  prettydoc::html_pretty:
    highlight: github
    theme: leonids
    toc: yes
subtitle: CUNY MSDS DATA 609
---

```{r, echo=FALSE, warning=FALSE, error=FALSE, cache=FALSE, results='hide', message=FALSE}

knitr::opts_chunk$set(echo = TRUE)
type <- 'latex'
```


```{r, echo=FALSE, warning=FALSE, error=FALSE, cache=FALSE, results='hide', message=FALSE}

library(knitr)
library(data.table)
library(xtable)
library(plotly)

```

# Problems

The below problems are taken from the text book:

A First Course in Mathematical Modeling, 5th Edition. Frank R. Giordano, William P. Fox, Steven B. Horton. ISBN-13: 9781285050904.

## Exercise \#12 Page Page 69.

A company with a fleet of trucks faces increasing maintenance costs as the age and mileage of the trucks increase.

### Solution

**Question:** What will be the total maintenance cost for a truck? 

**List the variables that affect the behavior you have identified.**

The variables that affect the behavior from the given sentence are:

[MAINTENANCE_COST, YEAR_BUILD, TRUCK_MILAGE]

Also, we could consider some other variables that could affect the maintenance costs, such as:

[DRIVER_EXPERTISE, TRAFFIC_CONDITIONS, ROAD_CONDITIONS, ROUTE_HAS_STEEP_HILLS, TRUCK_TYPE, TRUCK_WEIGHT, NUMBER_OF_GEARS, NUMBER_OF_AXES, OIL_CHANGE_FREQUENCY, BREAKING_FREQUENCY, TIRE_CONDITION, N_DAYS_FROM_LAST_MAINTENANCE, MAINTENANCE_PARTS_COST, MAINTENANCE_JOB_DESCRIPTION, MAINTENANCE_LABOR_COST, N_HOURS_MAINTENANCE_LABOR]

**Which variables will be neglected completely?**

I consider that the following variables will be neglected.

[DRIVER_GENDER, DRIVER_AGE, MAX_LOAD]

**Which might be considered as constants initially?**

The variables that I would consider as constants in an initial form will be:

[DRIVER_EXPERTISE, TRAFFIC_CONDITIONS, ROAD_CONDITIONS, OIL_CHANGE_FREQUENCY, MAINTENANCE_LABOR_COST, N_HOURS_MAINTENANCE_LABOR]

**Can you identify any submodels you would want to study in detail?**

I believe that we could construct some sub models as follows:

BREAKING_FREQUENCY = f(ROAD_CONDITIONS, ROUTE_HAS_STEEP_HILLS, NUMBER_OF_GEARS)

OIL_CHANGE_FREQUENCY = f(YEAR_BUILD, TRUCK_MILAGE, NUMBER_OF_GEARS)

TIRE_CONDITION = f(DRIVER_EXPERTISE, BREAKING_FREQUENCY, ROAD_CONDITIONS, ROUTE_HAS_STEEP_HILLS, NUMBER_OF_AXES)

MAINTENANCE_LABOR_COST = f(N_HOURS_MAINTENANCE_LABOR)

**Identify any data you would want collected.**

I would like to collect:

[YEAR_BUILD, TRUCK_MILAGE, DRIVER_EXPERTISE, TRAFFIC_CONDITIONS, ROAD_CONDITIONS, ROUTE_HAS_STEEP_HILLS, TRUCK_TYPE, TRUCK_WEIGHT, NUMBER_OF_GEARS, NUMBER_OF_AXES, N_DAYS_FROM_LAST_MAINTENANCE, MAINTENANCE_PARTS_COST, MAINTENANCE_JOB_DESCRIPTION, MAINTENANCE_LABOR_COST, N_HOURS_MAINTENANCE_LABOR]

## Exercise \#11 Page Page 79.

Determine whether the data set supports the stated proportionality model.

$y \propto x^3$

```{r, echo=FALSE}
y <- c(0, 1, 2, 6, 14, 24, 37, 58, 82, 114)
x <- c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10)

t.df <- data.frame(y, x)
```

```{r, echo=FALSE, results='asis'}
print(xtable(t.df, caption='', digits=0), hline.after=c(0,10), include.rownames=FALSE, type=type, comment=FALSE)
```

### Solution

In order to support a proportionality model; it must satisfy as follows:

$y = kx^3$ for a constant $k$.

Let's plot our given data as it is:

```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE, results='hide'}
pdf <- plot(t.df$y~t.df$x,
            type="p", 
            col="blue",
            main='Given data',
            xlab="x", 
            ylab="y")
pdf
```

Now, let's calculate $x^3$.

```{r, echo=FALSE}
t.df$`x^3` <- (t.df$x)^3
```

```{r, echo=FALSE, results='asis'}
print(xtable(t.df, caption='', digits=0), hline.after=c(0,10), include.rownames=FALSE, type=type, comment=FALSE)
```

Now, let's plot again and see it's behavior.


```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE, results='hide'}
pdf <- plot(t.df$y~t.df$`x^3`,
            type="p", 
            col="blue",
            main='Given data but with x^3',
            xlab="x", 
            ylab="y")
pdf
```

As we can see, there now seems to be some sort of linearity; hence we could calculate our slope as follows:

$k = \frac{114 - 0}{1000 - 1} = \frac{114}{999}$

From here, we could deduce that our model will be something similar to:

$y = \frac{114}{999} x^3$

Let's plot our linear model and see if it actually follows the points.

```{r, echo=FALSE}
plot(t.df$y~t.df$`x^3`,
            type="p", 
            col="blue",
            main='Given data but with x^3',
            xlab="x", 
            ylab="y")
lines(t.df$`x^3`, 114/999 * t.df$`x^3`, col="red", type="l", pch=17)
legend("topleft", 
       c("y","linear model"),
       fill=c("blue","red")
       )
```

From the above visual it seems that our linear model follows all points very accurate.

Now, I will find the predicted $y$ value rounded to two decimals and compare them to our given data in order to obtain an Error difference.

```{r, echo=FALSE}
t.df$`predicted y` <- round(114/999 * t.df$`x^3`,2)
t.df$Error <- t.df$`predicted y` - t.df$y
```

```{r, echo=FALSE, results='asis'}
print(xtable(t.df, caption='Error shows no decimals, but calculations made to two decimals.', digits=0), hline.after=c(0,10), include.rownames=FALSE, type=type, comment=FALSE)
```

```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE, results='hide'}
plot(t.df$Error~t.df$`x^3`,
            type="p", 
            col="red",
            main='Error difference',
            xlab="x", 
            ylab="predicted 'y' - y")
abline(h=0, col='blue')
```


```{r, echo=FALSE}
max_error = max(abs(t.df$Error))
```

From the above graph, we can find out that our maximum error difference is about `r max_error`.

# END.