---
title: "Lab 8 - Multiple linear regression"
author: "Duubar Villalobos Jimenez   mydvtech@gmail.com"
date: "May 7, 2017"
output:
  pdf_document: default
  html_document: default
  prettydoc::html_pretty:
    highlight: github
    theme: leonids
subtitle: CUNY MSDA DATA 606
---

```{r, message=FALSE, results='hide'}
library(knitr)
library(DATA606)
```


## Grading the professor

Many college courses conclude by giving students the opportunity to evaluate 
the course and the instructor anonymously. However, the use of these student 
evaluations as an indicator of course quality and teaching effectiveness is 
often criticized because these measures may reflect the influence of 
non-teaching related characteristics, such as the physical appearance of the 
instructor. The article titled, "Beauty in the classroom: instructors' 
pulchritude and putative pedagogical productivity" (Hamermesh and Parker, 2005) 
found that instructors who are viewed to be better looking receive higher 
instructional ratings. (Daniel S. Hamermesh, Amy Parker, Beauty in the 
classroom: instructors pulchritude and  putative pedagogical productivity, 
*Economics of Education Review*, Volume 24, Issue 4, August 2005, Pages 369-376, 
ISSN 0272-7757, 10.1016/j.econedurev.2004.07.013. [http://www.sciencedirect.com/science/article/pii/S0272775704001165](http://www.sciencedirect.com/science/article/pii/S0272775704001165).)

In this lab we will analyze the data from this study in order to learn what goes 
into a positive professor evaluation.

## The data

The data were gathered from end of semester student evaluations for a large 
sample of professors from the University of Texas at Austin. In addition, six 
students rated the professors' physical appearance. (This is slightly modified 
version of the original data set that was released as part of the replication 
data for *Data Analysis Using Regression and Multilevel/Hierarchical Models* 
(Gelman and Hill, 2007).) The result is a data frame where each row contains a 
different course and columns represent variables about the courses and professors.

```{r load-data, eval=TRUE}
load("more/evals.RData")
```

variable         | description
---------------- | -----------
`score`          | average professor evaluation score: (1) very unsatisfactory - (5) excellent.
`rank`           | rank of professor: teaching, tenure track, tenured.
`ethnicity`      | ethnicity of professor: not minority, minority.
`gender`         | gender of professor: female, male.
`language`       | language of school where professor received education: english or non-english.
`age`            | age of professor.
`cls_perc_eval`  | percent of students in class who completed evaluation.
`cls_did_eval`   | number of students in class who completed evaluation.
`cls_students`   | total number of students in class.
`cls_level`      | class level: lower, upper.
`cls_profs`      | number of professors teaching sections in course in sample: single, multiple.
`cls_credits`    | number of credits of class: one credit (lab, PE, etc.), multi credit.
`bty_f1lower`    | beauty rating of professor from lower level female: (1) lowest - (10) highest.
`bty_f1upper`    | beauty rating of professor from upper level female: (1) lowest - (10) highest.
`bty_f2upper`    | beauty rating of professor from second upper level female: (1) lowest - (10) highest.
`bty_m1lower`    | beauty rating of professor from lower level male: (1) lowest - (10) highest.
`bty_m1upper`    | beauty rating of professor from upper level male: (1) lowest - (10) highest.
`bty_m2upper`    | beauty rating of professor from second upper level male: (1) lowest - (10) highest.
`bty_avg`        | average beauty rating of professor.
`pic_outfit`     | outfit of professor in picture: not formal, formal.
`pic_color`      | color of professor's picture: color, black & white.

## Exploring the data

1.  Is this an observational study or an experiment? The original research
    question posed in the paper is whether beauty leads directly to the
    differences in course evaluations. Given the study design, is it possible to
    answer this question as it is phrased? If not, rephrase the question.

**Answer:**

This is an observational study. Experiments provide control and experimental groups and there's none in this case.

In an observational study we cannot set causation between the explanatory and response variables; but to find a correlation instead.

A good question could be: Does the instructor's physical appearance has impact to student course evaluation?

2.  Describe the distribution of `score`. Is the distribution skewed? What does 
    that tell you about how students rate courses? Is this what you expected to 
    see? Why, or why not?

**Answer:**

```{r}
hist(evals$score)
```

- Is the distribution skewed? Yes, the evaluation scores are `left skewed`.

- What does that tell you about how students rate courses? 

There are more positive physical appearance evaluations than negative for their teachers based on their students response.

- Is this what you expected to see? Why, or why not?

To be honest, I was not expecting any specific response. But from classroom experience, I would say that knowledge and kindness beauty transpose physical appearance providing extra factors to consider.

3.  Excluding `score`, select two other variables and describe their relationship 
    using an appropriate visualization (scatterplot, side-by-side boxplots, or 
    mosaic plot).
    
**Answer:**

For this, I have selected `age` and `bty_avg`.

```{r}
plot(x = evals$age, y = evals$bty_avg)
```


Based on pure visualization, in the scatter plot there seems to be a negative relationship in between the two variables as the instructor gets older.


```{r}
boxplot(evals$bty_avg ~ evals$age)
```

By observing the boxplot, we noticed that some ages have a wide `bty_avg` range while others have a low range; in some cases we can notice some outliers on some of the grading ages.

```{r}
mosaicplot(~ age + bty_avg , data = evals, color = TRUE)
```

By observing this mosaic plot, we can visually identify a few grading observations or when age groups were the most and less graded. 

## Simple linear regression

The fundamental phenomenon suggested by the study is that better looking teachers
are evaluated more favorably. Let's create a scatterplot to see if this appears 
to be the case:

```{r scatter-score-bty_avg, eval = FALSE}
plot(evals$score ~ evals$bty_avg)
```


Before we draw conclusions about the trend, compare the number of observations 
in the data frame with the approximate number of points on the scatterplot. 
Is anything awry?

**Answer:**

```{r}
nrow(evals)
```

The number of observations are $463$; however there seems to be less than 463 points in the scattered plot.

4.  Replot the scatterplot, but this time use the function `jitter()` on the
    $y$- or the $x$-coordinate. (Use `?jitter` to learn more.) What was
    misleading about the initial scatterplot?

**Answer:**    
    
```{r}
scoreNoise <- jitter(evals$score, factor = 1, amount = NULL)
plot(scoreNoise ~ evals$bty_avg)
```

It seems that the original scattered plot was plotting only one point representing the score mean of the points that were very near to each other; providing a misrepresentation of the whole scattered plot since many point were hidden.


5.  Let's see if the apparent trend in the plot is something more than
    natural variation. Fit a linear model called `m_bty` to predict average
    professor score by average beauty rating and add the line to your plot
    using `abline(m_bty)`. Write out the equation for the linear model and
    interpret the slope. Is average beauty score a statistically significant
    predictor? Does it appear to be a practically significant predictor?

**Answer:**

```{r}
m_bty <- lm(evals$score ~ evals$bty_avg)
plot(scoreNoise ~ evals$bty_avg)
abline(m_bty)
```

**Find Correlation**

```{r}
scorNbty_Cor <- cor(evals$score, evals$bty_avg)
```

The correlation in between the `score` and `bty_avg` is: `r scorNbty_Cor`.

```{r}
summary(m_bty)
```

The equation for the linear model is: $\hat{y} = 3.88034 + 0.06664 \cdot bty\_avg$

- Is average beauty score a statistically significant predictor?

Yes, `bty_avg` is a statistically significant predictor of evaluation score with p-value close of $0$.

- Does it appear to be a practically significant predictor?

It appears not to be a practically significant predictor of evaluation score since for every $1$ point increase in `bty_ave`, the model only predicts an increase of $0.06664$. this is not a very significant change in the evaluation score, hence is not a significant predictor.

6.  Use residual plots to evaluate whether the conditions of least squares
    regression are reasonable. Provide plots and comments for each one (see
    the Simple Regression Lab for a reminder of how to make these).
    
**Answer:**

```{r}
plot_ss(x = evals$bty_avg, y = scoreNoise, showSquares = TRUE)
```


When fitting a least squares line, we generally require:

`Linearity`: *"The data should show a linear trend"*

```{r}
plot(m_bty$residuals ~ evals$bty_avg)
abline(h = 0)
```

The relationship looks linear. By looking at the residual plot as the variability of residuals is approximately constant across the distribution but does not indicate any curvatures or any indication of non-normality.

`Nearly normal residuals`: *"Generally the residuals must be nearly normal. When this condition is found to be unreasonable, it is usually because of outliers or concerns about influential points"*.

```{r}
hist(m_bty$residuals)
```

By looking at the histogram we can observe that the residuals do not follow some sort of normality in respect to their frequency distribution. In this case, this condition is **NOT** met.

```{r}
qqnorm(m_bty$residuals)
qqline(m_bty$residuals)
```

And by looking at the Q-Q Plot, we can observe how the distribution is **NOT** following around a straight line. hence we can conclude that this model does not satisfies the nearly normal residuals condition.

`Constant variability`: *"The variability of points around the least squares line remains roughly constant"*. 
Based on the plot the variability of points around the least squares line remains roughly constant so the condition constant variability has been met.

`Independent observations`: We can considered this to be met since is this observational study represents less than 10% of the population. 

Since one of the conditions of least squares regression is not met; we conclude that this regression method is not satisfactory for considering `avg_bty` as a predictor of `score`.

## Multiple linear regression

The data set contains several variables on the beauty score of the professor: 
individual ratings from each of the six students who were asked to score the 
physical appearance of the professors and the average of these six scores. Let's 
take a look at the relationship between one of these scores and the average 
beauty score.

```{r bty-rel, eval = FALSE}
plot(evals$bty_avg ~ evals$bty_f1lower)
cor(evals$bty_avg, evals$bty_f1lower)
```

As expected the relationship is quite strong - after all, the average score is 
calculated using the individual scores. We can actually take a look at the 
relationships between all beauty variables (columns 13 through 19) using the 
following command:

```{r bty-rels, eval = FALSE}
plot(evals[,13:19])
```

These variables are collinear (correlated), and adding more than one of these 
variables to the model would not add much value to the model. In this 
application and with these highly-correlated predictors, it is reasonable to use
the average beauty score as the single representative of these variables.

In order to see if beauty is still a significant predictor of professor score 
after we've accounted for the gender of the professor, we can add the gender 
term into the model.

```{r scatter-score-bty_avg_gender}
m_bty_gen <- lm(score ~ bty_avg + gender, data = evals)
summary(m_bty_gen)
```

7.  P-values and parameter estimates should only be trusted if the
    conditions for the regression are reasonable. Verify that the conditions
    for this model are reasonable using diagnostic plots.
    
**Answer:**

```{r, message=FALSE, warning=FALSE, error=FALSE, message=FALSE, results='hide'}
plot(scoreNoise ~ evals$bty_avg)
abline(m_bty_gen)
```

```{r}
plot(scoreNoise ~ evals$gender)
```

When fitting a least squares line, we generally require:

`Nearly normal residuals`: *"Generally the residuals must be nearly normal. When this condition is found to be unreasonable, it is usually because of outliers or concerns about influential points"*.

```{r}
hist(m_bty_gen$residuals)
```

By looking at the histogram we can observe that the residuals seems not to follow some sort of normality in respect to their frequency distribution. 

```{r}
qqnorm(m_bty_gen$residuals)
qqline(m_bty_gen$residuals)
```

And by looking at the Q-Q Plot, we can observe how the distribution tends to follow around a straight line but then it deviates due to outliers.

`Constant variability`: *"The variability of points around the least squares line remains roughly constant"*. 
Based on the plot the variability of points around the least squares line remains roughly constant so the condition constant variability has been met.

```{r}
plot(m_bty_gen$residuals ~ evals$bty_avg)
abline(h = 0)
```


`Independent observations`: We can considered this to be met since is this observational study represents less than 10% of the population.


8.  Is `bty_avg` still a significant predictor of `score`? Has the addition
    of `gender` to the model changed the parameter estimate for `bty_avg`?

**Answer:** Yes, `bty_avg` is a significant predictor. With the addition of `gender` it has added even more significance since the p-value became smaller.

Note that the estimate for `gender` is now called `gendermale`. You'll see this 
name change whenever you introduce a categorical variable. The reason is that R 
recodes `gender` from having the values of `female` and `male` to being an 
indicator variable called `gendermale` that takes a value of $0$ for females and
a value of $1$ for males. (Such variables are often referred to as "dummy" 
variables.)

As a result, for females, the parameter estimate is multiplied by zero, leaving 
the intercept and slope form familiar from simple regression.

\[
  \begin{aligned}
\widehat{score} &= \hat{\beta}_0 + \hat{\beta}_1 \times bty\_avg + \hat{\beta}_2 \times (0) \\
&= \hat{\beta}_0 + \hat{\beta}_1 \times bty\_avg\end{aligned}
\]

We can plot this line and the line corresponding to males with the following 
custom function.

```{r twoLines}
multiLines(m_bty_gen)
```


9.  What is the equation of the line corresponding to males? (*Hint:* For males, the parameter estimate is multiplied by 1.) For two professors who received the same beauty rating, which gender tends to have the higher course evaluation score?

**Answer:**
    
```{r}
summary(m_bty_gen)
```

Since the gender `male` can be represented with $1$. From the original equation we have as follows:

$\widehat{score} = \hat{\beta}_0 + \hat{\beta}_1 \times bty\_avg + \hat{\beta}_2 \times Male$

$\widehat{score} = 3.74734 + 0.07416 \times bty\_avg + 0.17239 \times 1$

$\widehat{score} = 3.91973 + 0.07416 \times bty\_avg$

For two professors who received the same beauty rating, which gender tends to have the higher course evaluation score?

In this predictive model, Male professors will receive the highest score by 0.17239.

The decision to call the indicator variable `gendermale` instead of`genderfemale`
has no deeper meaning. R simply codes the category that comes first 
alphabetically as a $0$. (You can change the reference level of a categorical 
variable, which is the level that is coded as a 0, using the`relevel` function. 
Use `?relevel` to learn more.)

10. Create a new model called `m_bty_rank` with `gender` removed and `rank` 
    added in. How does R appear to handle categorical variables that have more 
    than two levels? Note that the rank variable has three levels: `teaching`, 
    `tenure track`, `tenured`.

```{r}
m_bty_rank <- lm(score ~ bty_avg + rank, data = evals)
summary(m_bty_rank)
```

The interpretation of the coefficients in multiple regression is slightly 
different from that of simple regression. The estimate for `bty_avg` reflects
how much higher a group of professors is expected to score if they have a beauty
rating that is one point higher *while holding all other variables constant*. In
this case, that translates into considering only professors of the same rank 
with `bty_avg` scores that are one point apart.

## The search for the best model

We will start with a full model that predicts professor score based on rank, 
ethnicity, gender, language of the university where they got their degree, age, 
proportion of students that filled out evaluations, class size, course level, 
number of professors, number of credits, average beauty rating, outfit, and 
picture color.

11. Which variable would you expect to have the highest p-value in this model? 
    Why? *Hint:* Think about which variable would you expect to not have any 
    association with the professor score.

**Answer:** Language; since we are evaluating the physical appearance of the instructor; the language should not have a major association with the professor score.

Let's run the model...

```{r m_full, tidy = FALSE}
m_full <- lm(score ~ rank + ethnicity + gender + language + age + cls_perc_eval 
             + cls_students + cls_level + cls_profs + cls_credits + bty_avg 
             + pic_outfit + pic_color, data = evals)
summary(m_full)
```

12. Check your suspicions from the previous exercise. Include the model output
    in your response.
    
**Answer:**

The highest p value for this model is $0.77806$ for `cls_profs`.

The p value for the variable I was expecting to be the highest is $0.03965$ for `language`; which is a big difference in between the two of them.

13. Interpret the coefficient associated with the ethnicity variable.

**Answer:**

By considering all other variables being equal; the score for instructors that are **not minority** tends to be $0.1234929$ higher.

14. Drop the variable with the highest p-value and re-fit the model. Did the
    coefficients and significance of the other explanatory variables change?
    (One of the things that makes multiple regression interesting is that
    coefficient estimates depend on the other variables that are included in
    the model.) If not, what does this say about whether or not the dropped
    variable was collinear with the other explanatory variables?

**Answer:**


```{r m_full1, tidy = FALSE}
m_full1 <- lm(score ~ rank + ethnicity + gender + language + age + cls_perc_eval 
             + cls_students + cls_level + cls_credits + bty_avg 
             + pic_outfit + pic_color, data = evals)
summary(m_full1)
```

```{r}
m_full$coefficients == m_full1$coefficients
```

Did the coefficients and significance of the other explanatory variables change?

**Yes**, the coefficients changed, which means the dropped variable depends on other variables as well.

15. Using backward-selection and p-value as the selection criterion,
    determine the best model. You do not need to show all steps in your
    answer, just the output for the final model. Also, write out the linear
    model for predicting score based on the final model you settle on.

**Answer:**

```{r}
m_full2 <- lm(score ~ gender + language + age + cls_perc_eval 
             + cls_credits + bty_avg + pic_color, data = evals)
summary(m_full2)
```

```{r,}
score <- function(gender, language, age, cls_perc_eval, cls_credits, bty_avg, pic_color){
    
score <-( 3.967255 
        + 0.221457 * gender
        - 0.281933 * language
        - 0.005877 * age
        + 0.004295 * cls_perc_eval
        + 0.444392 * cls_credits
        + 0.048679 * bty_avg
        - 0.216556 * pic_color)

return(round(score,1))

}

backwardSelection <- score(1, 1, evals$age, evals$cls_perc_eval, 1, evals$bty_avg, 1)

compareScores <- data.frame(evals$score, backwardSelection, backwardSelection - evals$score)
names(compareScores) <- c("Original", "Prediction", "Difference")
```

```{r}
kable(head(compareScores,20))
```


16. Verify that the conditions for this model are reasonable using diagnostic 
    plots.

**Answer:**

- By looking at the histogram:

```{r}
hist(m_full2$residuals) 
```

- Normal Probability Plot

```{r}
qqnorm(m_full2$residuals)
qqline(m_full2$residuals)
```

- Gender

```{r}
plot(scoreNoise ~ evals$gender)
```

- Language

```{r}
plot(scoreNoise ~ evals$language)
```

- Age 

```{r}
plot(scoreNoise ~ evals$age) 
```

- cls_perc_eval

```{r}
plot(scoreNoise ~ evals$cls_perc_eval) 
```

- Class Credits

```{r}
plot(scoreNoise ~ evals$cls_credits)
```

- bty_avg

```{r}
plot(scoreNoise ~ evals$bty_avg) 
```

- Pic Color

```{r}
plot(scoreNoise ~ evals$pic_color)
```

The variables above are related to the score.

17. The original paper describes how these data were gathered by taking a
    sample of professors from the University of Texas at Austin and including 
    all courses that they have taught. Considering that each row represents a 
    course, could this new information have an impact on any of the conditions 
    of linear regression?

**Answer:**

From perspective, class courses are independent of each other. By having this condition of independence, evaluation scores from one course is independent of the other. If an instructor teaches more than one course it should not affect, however if the same student takes two or more classes with the same instructor this will affect the outcome since independence will not be satisfied.

18. Based on your final model, describe the characteristics of a professor and 
    course at University of Texas at Austin that would be associated with a high
    evaluation score.
    
**Answer:**

Based on this model, the characteristics of the highest scores will be obtained by male instructors who obtained their degree in an english speaking university, teach one credit class and has a black and white picture.


19. Would you be comfortable generalizing your conclusions to apply to professors
    generally (at any university)? Why or why not?
    
**Answer:**

No, this report was not conducted as an experiment but based on an observational study in one given university. Also, the definition of beauty has changed over time, cultural values also change, these results may be different in other geographically different university or in a different time frame.

<div id="license">
This is a product of OpenIntro that is released under a [Creative Commons Attribution-ShareAlike 3.0 Unported](http://creativecommons.org/licenses/by-sa/3.0). This lab was written by 
Mine &Ccedil;etinkaya-Rundel and Andrew Bray.
</div>
