---
title: "Homework 3.2"
author: "Duubar Villalobos Jimenez   mydvtech@gmail.com"
date: "October 07, 2018"
output:
  pdf_document: default
  html_document: default
  prettydoc::html_pretty:
    highlight: github;l'
    theme: leonids
    toc: yes
subtitle: CUNY MSDS DATA 609
---

```{r, echo=FALSE, warning=FALSE, error=FALSE, cache=FALSE, results='hide', message=FALSE}

knitr::opts_chunk$set(echo = TRUE)

```


```{r, echo=FALSE, warning=FALSE, error=FALSE, cache=FALSE, results='hide', message=FALSE}

library(knitr)

```

# Problems

The below problems are taken from the text book:

A First Course in Mathematical Modeling, 5th Edition. Frank R. Giordano, William P. Fox, Steven B. Horton. ISBN-13: 9781285050904.

## Exercise \#6 Page 264

Solve using graphical analysis.

Maximize $10x + 35y$

subject to

$8x + 6y \le 48$ (board-feet of lumber)

$4x + y  \le 20$ (hour of carpentry)

$y \ge 5$ (demand)

$x,y \ge 0$ (non-negativity)


### Solution


```{r, echo=FALSE}
f1 <- function(x){(48 - 8*x)/6}
f2 <- function(x){20 - 4*x}
f <- function(x,y){10*x + 35*y}

x <- c(-1, 6)
y <- c(-1, 25)

plot(f1, 
     type='h',
     ylab = 'y',
     xlab = 'x',
     xlim=x,
     ylim=y,
     col='red')
abline(v=0)
plot(f2, 
     type='h',
     ylab = 'y',
     xlab = 'x',
     xlim=x,  
     ylim=y,
     col='blue',
     add=TRUE)
abline(h=5, col='green')
polygon(c(0,0,18/8),c(5,8,5),col="gray")
lnames <- c('8x + 6y = 48', '4x + y = 20', 'y =5')
legend('topright', lnames, col = c('red','blue', 'green'), lty = 1) 
```

From the above graph, we can see that we have 3 main points in the y axis as follows $A(0,8)$, $B(0,20)$, $E(0,5)$ and then two more intersections with $y=5$ as follows $C(\frac{15}{4},5)$ and $D(\frac{18}{8},5)$ and by evaluating such points on the objective function, we obtain the following results:

```{r}
x <- c(0,0,18/8)
y <- c(48/6,5,5)
f.df <- data.frame(x,y) 
f.df$`10x + 35y` <- 10*f.df$x + 35*f.df$y

f.df
```

From the above table, after evaluating our critic points in $10x + 35y$, we found out that the maximum value is 280.

```{r, echo=FALSE}
fmax <- function(x){(280 - 10*x)/35}

x <- c(-1, 5)
y <- c(-1, 22)

plot(f1, 
     type='l',
     ylab = 'y',
     xlab = 'x',
     xlim=x,
     ylim=y)
abline(v=0)
plot(f2, 
     type='l',
     ylab = 'y',
     xlab = 'x',
     xlim=x,  
     ylim=y,
     add=TRUE)
abline(h=5)
plot(fmax, 
     type='S',
     ylab = 'y',
     xlab = 'x',
     xlim=x,  
     ylim=y,
     col='red',
     add=TRUE)
polygon(c(0,0,18/8),c(5,8,5),col="gray")
text(x = 0.1, 
     y = 8.5,
     labels = 'A(0,8)',
     adj = 0)
text(x = 0.1, 
     y = 20.5,
     labels = 'B(0,20)',
     adj = 0)
text(x = 3.85, 
     y = 5.5,
     labels = 'C(3.75,5)',
     adj = 0)
text(x = 2.35, 
     y = 5.5,
     labels = 'D(2.25,5)',
     adj = 0)
text(x = 0.1, 
     y = 5.5,
     labels = 'E(0,5)',
     adj = 0)
text(x = 3.855, 
     y = 9.5,
     labels = '10x + 35y = 280',
     col='red',
     adj = 0)
```

## Exercise \#6 Page 264

Solve using the method of Algebraic Solutions.

Maximize $10x + 35y$

subject to

$8x + 6y \le 48$ (board-feet of lumber)

$4x + y  \le 20$ (hour of carpentry)

$y \ge 5$ (demand)

$x,y \ge 0$ (non-negativity)


### Solution

First, we convert the inequalities to equations by adding new necessary non negative "slack" variables $z_1$, $z_2$ and $z_3$. If either $z_1$, $z_2$ or $z_3$ is negative, then the constrain is not satisfied. Thus the problem becomes:

Maximize $10x_1 + 35y_1$

subject to

$8x_1 + 6y_1 + z_1 = 48$

$4x_1 + y_1 + z_2 = 20$

$y_1 + z_3 = 5$

$x_1, y_2, z_1, z_2, z_3 \ge 0$ (non-negativity)

We now consider the entire set of 5 variables ${x_1, y_1, z_1, z_2, z_3}$ to determine the possible intersection point in the $x_1 x_2$ plane.

Let's begin by assigning the variables $x_1$ and $x_2$ the value of zero; resulting in the following set of equations:

$z_1 = 48$

$z_2 = 20$ 

Which is could be a feasible intersection point $A(0,0)$ because all four variables are non negative; but the second constrain is violated by 5 units, indicating that the intersection point $A(0,0)$ is infeseable.

For the second intersection point we choose the variable $x_1$ and $z_1$ and set them to zero, yielding the system:

$6y_1 = 48$

$y_1 + z_2 = 20$


That has solution $y_1 = 8$ and $z_2 = 12$, which is feasible intersection point $B(0,8)$.

For the third intersection point we choose $x_1$ and $z_2$ and set them to zero, yielding the system:

$6y_1 + z_1 = 48$

$y_1 = 20$

That has solution $y_1 = 20$ and $z_1 = -72$. Thus, the first constrain is violated by 72 units and the second constrain is violated by 77 units, indicating that the intersection point $C(0,20)$ is infeseable.

In a similar fashion choosing $z_1$ and $z_2$ and setting to zero, yielding the system:

$8x_1 + 6y_1 = 48$

$4x_1 + y_1 = 20$

That has solution $x_1 = \frac{9}{2}$ and $y_1 = 2$ Thus, the second constrain is violated by 3 units, indicating that the intersection point $D(\frac{9}{2},2)$ is infeseable.

For the fourth intersection point we choose $y_1 = 5$ and $z_1 = 0$, yielding the system:

$8x_1 + 30 = 48$

$5 + z_3 = 5$

That has solution $x_1 = \frac{18}{8}$ and $z_3 = 0$ which is feasible intersection point $E( \frac{18}{8},5)$.

For the fifth intersection point we choose $y_1 = 5$ and $z_2 = 0$, yielding the system:

$4x_1 + 5 = 20$

$5 + z_3 = 5$

That has solution $x_1 = \frac{15}{4}$ and $z_3 = 0$ which is feasible intersection point $F( \frac{15}{4},5)$.

For the sixth intersection point we choose, $x_1 = 0$ and $y_1 = 5$, yielding the system:

$30 + z_1 = 48$

$5 + z_2 = 20$

$z_3 = 5$

That has solution $z_1 = 18$, $z_2 = 15$ and $z_3 = 5$ which is feasible intersection point $G(0,5)$.

In summary, of the six possible intersection points in the $x_1 x_2$ plane, the following were found to be feasible.

$B(0,8)$

$E( \frac{18}{8}, 5)$

$F( \frac{15}{4}, 5)$

$G(0,5)$


```{r}
x <- c(0,18/8, 15/4, 0)
y <- c(8,5,5,5)
f.df <- data.frame(x,y) 
f.df$`10x + 35y` <- 10*f.df$x + 35*f.df$y

f.df
```

**Answer** The maximum generated will be with $x=0$ and $y=8$.

#  End.

