---
title: "Homework Week 2"
author: "Duubar Villalobos Jimenez   mydvtech@gmail.com"
date: "February 12, 2017"
output:
  html_document: default
  pdf_document: default
subtitle: CUNY MSDA DATA 607
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(htmlTable)
library(RMySQL)
library(sqldf)
```

Assignment - SQL a nd R  

Choose six recent popular movies.  Ask at least five people that you know (friends,  family,  classmates, imaginary friends) to rate each of these movie that they have seen on a scale of 1 to 5.  Take the results (observations) and store them in a SQL   database.  Load the information into an R dataframe. Your deliverables should include your SQL   scripts and your R Markdown code,  posted to GitHub. This is by design a very open ended assignment.  A variety of reasonable approaches are acceptable.  You can (and should) blank out your SQL   password if your solution requires it; otherwise,  full credit requires that your code is "reproducible,"  with the assumption that I have the same database server and R software. You may work in a small group.   


## Solution

### Need to type Local MySQL Root password
```{r}
# Need to type root password for the local database
myLocalPassword <- 'pswrd'
```

### Connect to database
```{r}
# Connect to database
my.database = dbConnect(MySQL(), user='root', password = myLocalPassword, dbname='villalobos-movies', host='localhost')
dbListTables(my.database)
```

### Display Users Table
```{r}
# -- ----------------------------------------------
# -- Table `users`
# -- ----------------------------------------------
my.users <- dbSendQuery(my.database, "SELECT * FROM tblUsers;")
dbFetch(my.users)
```

### Display Movies Table
```{r}
# -- ----------------------------------------------
# -- Table `movies`
# -- ----------------------------------------------
my.movies <- dbSendQuery(my.database, "SELECT * FROM tblmovies;")
dbFetch(my.movies)
```


### Display Reviews Table Raw info
```{r}
# -- ----------------------------------------------
# -- Table `reviewss`
# -- ----------------------------------------------
my.reviews <- dbSendQuery(my.database, "SELECT * FROM tblreviews;")
dbFetch(my.reviews)
```


### Read data from MySQL all tables combined into one query
```{r}
movie.data = dbSendQuery(my.database, "SELECT 
M.title As 'Title',
M.lenght AS 'Lenght',
U.fname As 'User',
R.rating As 'Rating',
R.review AS 'Review'
FROM tblMovies AS M
JOIN tblReviews AS R
ON M.movie_id = R.movie_id
JOIN tblUsers AS U
ON U.user_id = R.user_id;")
dbFetch(movie.data)
```

### Only one user who gave reviews
```{r}
movie.dataSingleUser = dbSendQuery(my.database, "SELECT 
M.title As 'Title',
M.lenght AS 'Lenght',
U.fname As 'User',
R.rating As 'Rating',
R.review AS 'Review'
FROM tblMovies AS M
JOIN tblReviews AS R
ON M.movie_id = R.movie_id
JOIN tblUsers AS U
ON U.user_id = R.user_id
WHERE U.user_id = 1;")
dbFetch(movie.dataSingleUser)
```

### Only one Rating score
```{r}
movie.dataRating = dbSendQuery(my.database, "SELECT 
M.title As 'Title',
M.lenght AS 'Lenght',
U.fname As 'User',
R.rating As 'Rating',
R.review AS 'Review'
FROM tblMovies AS M
JOIN tblReviews AS R
ON M.movie_id = R.movie_id
JOIN tblUsers AS U
ON U.user_id = R.user_id
WHERE R.rating = 5;")
dbFetch(movie.dataRating)
```

### Database disconnect
```{r}
### Disconnect
#dbGetQuery(my.database, "show processlist")
#dbGetQuery(my.database, "kill 14")
```

