---
title: "Homework Week 3"
author: "Duubar Villalobos Jimenez   mydvtech@gmail.com"
date: "February 19, 2017"
output:
  pdf_document: default
  html_document: default
subtitle: CUNY MSDA DATA 607
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Please deliver links to an R Markdown file (in GitHub and rpubs.com) with solutions to problems 3 and 4 from chapter 8 of Automated Data Collection in R.  Problem 9 is extra credit.


# Library definition
```{r library_setups, warning=FALSE, error=FALSE, cache=FALSE}
# Need to employ stringr for Regular Expressions
library(stringr)
```

# Problems

## 3. Copy the introductory example. The vector name stores the extracted names.

### Original Example
```{r}
raw.data <- "555-1239Moe Szyslak(636) 555-0113Burns, C. Montgomery555-6542Rev. Timothy Lovejoy555 8904Ned Flanders636-555-3226Simpson, Homer5553642Dr. Julius Hibbert"
```

### Extracting the vector name
```{r}
name <- unlist(str_extract_all(raw.data, "[[:alpha:]., ]{2,}"))
name
```

#### (a) Use the tools of this chapter to rearrange the vector so that all elements conform to the standard first_name last_name.

```{r, warning=FALSE}
# Original Name vector obtained by using Regular Expressions
name
# Split name for those names that have the last name first separated by a comma
split_name <- str_split(name, ",")
split_name

# Create a data frame to work from there
split_name <- data.frame(split_name)
# Display the new data frame, in order to vizualice the two rows, from here we can observe the difference in between rows when the last name was separated by a comma.
split_name

# Assing the "Last Names" Row
ln <- data.frame(split_name[1,])
# Assing the "First Names" Row
fn <- data.frame(split_name[2,])

# Compare rows and proceed to create the desired "Firt Name then Last name" output by employing rbind
split_name <- ifelse(fn == ln, ln , rbind(fn, ln))
split_name
```

### (b) Construct a logical vector indicating whether a character has a title (i.e.,Rev. and Dr.).

```{r}
# Defining Tiles Vector
title <- c("Rev.","Dr.")

# Find out if the title is part of the name
names_wtitle <- ifelse(str_detect(name,title) == TRUE, "YES", "NO")

# Creating a data frame to represent if title is part of the name
names_wtitle <- data.frame (name=name, title=names_wtitle)
names_wtitle

```

### (c) Construct a logical vector indicating whether a character has a second name.

```{r}
# Identify if the name has a Middle name. The Regular expressions are generally and initial followed by a period.
middle_name <- ifelse(str_detect(name,"[A-Z]\\.") == TRUE, "YES", "NO")
middle_name
# Report next to a name in a data frame.
names_wmname <- data.frame (name, middle=middle_name)
names_wmname
```

## 4. Describe the types of strings that conform to the following regular expressions and construct an example that is matched by the regular expression.

### (a) [0-9]+\\$

This one represent a one digit from [0-9] repeating multiple times to the right at the end of a expression with a dollar ($) sign at the end of the number.

```{r}
raw.vector <- c("This is my example where 1234567890$ is represented at the end of the vector; it will report [1234567890$] two times with a dollar sign at the end", "Another good example is 123$ but not $123")
unlist(str_extract_all(raw.vector, "[0-9]+\\$"))
```

### (b) \\b[a-z]{1,4}\\b

In this example it will display all the words that are surrounded by edges \\b on both sides and composed of four letters or less but with LOWER CASE only.

```{r}
raw.vector <- c("In this example it will display all the words that are surrounded by edges \\b on both sides and composed of four letters or less but with LOWER CASE only.")
unlist(str_extract_all(raw.vector, "\\b[a-z]{1,4}\\b"))
```

### (c) .*?\\.txt$

In this example it will display all the sentences that are composed of a period . followed by a word that has the asterisk or star that could be optional and then followed by the extension.txt, for example:

```{r}
raw.vector <- c("In this example it will display all the sentences that are composed of a period . followed by a word that has the asterisk or star that could be optional and then followed by the extension.txt, for example:", "this is good my.homework.txt", "my.homew*rk.txt is a very good example not included because is not at the end of the sentence.", "but this one is great my.homew*rk.txt")
unlist(str_extract_all(raw.vector, ".*?\\.txt$"))
```

### (d) \\d{2}/\\d{2}/\\d{4}

This is for dates composed of up to two digit month, two digit day and four digit year separated with the slash symbol.

```{r}
raw.vector <- c("in this example we will extract 02/04/2016 and 08/09/1977","Also we will not be able to obtain 4/15/1879 since it doest nt match the two digit month, neither 02/19/17")
unlist(str_extract_all(raw.vector, "\\d{2}/\\d{2}/\\d{4}"))
```

### (e) <(.+?)>.+?</\\1>

After doing some research, this represents the Vector containing strings with any type of HTML tag.
The back reference removes the outer HTML tags.

The \\1 is doing a recall of <(.+?)> at the end of the regular expression.

```{r}
raw.vector <- c("<!DOCTYPE html><html><body>Hello World</body></html></html>")
unlist(str_extract_all(raw.vector, "<(.+?)>.+?</\\1>"))
```

## 9. The following code hides a secret message. Crack it with R and regular expressions. Hint: Some of the characters are more revealing than others! The code snippet is also available in the materials at www.r-datacollection.com.

clcopCow1zmstc0d87wnkig7OvdicpNuggvhryn92Gjuwczi8hqrfpRxs5Aj5dwpn0Tanwo
Uwisdij7Lj8kpf03AT5Idr3coc0bt7yczjatOaootj55t3Nj3ne6c4Sfek.r1w1YwwojigO
d6vrfUrbz2.2bkAnbhzgv4R9i05zEcrop.wAgnb.SqoU65fPa1otfb7wEm24k6t3sR9zqe5
fy89n6Nd5t9kc4fE905gmc4Rgxo5nhDk!gr

```{r}
# Raw vector
raw.vector <- "clcopCow1zmstc0d87wnkig7OvdicpNuggvhryn92Gjuwczi8hqrfpRxs5Aj5dwpn0TanwoUwisdij7Lj8kpf03AT5Idr3coc0bt7yczjatOaootj55t3Nj3ne6c4Sfek.r1w1YwwojigOd6vrfUrbz2.2bkAnbhzgv4R9i05zEcrop.wAgnb.SqoU65fPa1otfb7wEm24k6t3sR9zqe5fy89n6Nd5t9kc4fE905gmc4Rgxo5nhDk!gr"
# I noticed there are some upper case letters and some periods in between, so I run the code for Alphanumeric of lenght 1 with an optional period.
hidden_message <- unlist(str_extract_all(raw.vector, "[[:upper:].?]{1,}"))
hidden_message
# Since the periods work as a separator, we can replace them for blank spaces and also we can put all toguether in one sentence.
hidden_message <- str_replace_all(paste(hidden_message, collapse = ''), "[.]", " ")
# Final message
hidden_message
```

