---
title: "Homework 01"
author: "Duubar Villalobos Jimenez   mydvtech@gmail.com"
date: "September 2, 2018"
output:
  pdf_document: default
  html_document: default
  prettydoc::html_pretty:
    highlight: github
    theme: leonids
    toc: yes
subtitle: CUNY MSDS DATA 609
---

```{r, echo=FALSE, warning=FALSE, error=FALSE, cache=FALSE, results='hide', message=FALSE}

knitr::opts_chunk$set(echo = TRUE)
type <- 'latex'
```


```{r, echo=FALSE, warning=FALSE, error=FALSE, cache=FALSE, results='hide', message=FALSE}

library(knitr)
library(data.table)
library(xtable)
library(plotly)

```

# Problems

The below problems are taken from the text book:

A First Course in Mathematical Modeling, 5th Edition. Frank R. Giordano, William P. Fox, Steven B. Horton. ISBN-13: 9781285050904.

## Exercise \#10 Page Page 8.

Your grandparents have an annuity. The value of the annuity increases each month by an automatic deposit of 1\% interest on the previous month's balance. Your grandparents withdraw \$1000 at the beginning of each month for living expenses. Currently, they have \$50,000 in the annuity. Model the annuity with a dynamical system. Will the annuity run out of money? When? Hint: What value will $a_n$ have when the annuity is depleted?

### Solution

Basically, the change in the amount increases by the amount paid in the form of interest and decreases by the amount withdraw each month.

Change = $\Delta a_n = a_{n+1} - a_n = 0.01 \cdot a_n - 1000$

The dynamical system will be:

$a_0 = 50000$

$a_{n+1} = a_n + 0.01 \cdot a_n - 1000$

The below table represent the values for the upcoming months.

```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
n <- 0
Balance <- 50000
annuity_Table <- data.frame(`Month` = n, `Balance` = Balance)
```

```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
new_Balance <- Balance
while (new_Balance >= 0) {
  
  n <- n + 1
  new_Balance <- round(annuity_Table$Balance[n] + 0.01 * annuity_Table$Balance[n] - 1000,2)
  new_row <- data.frame(`Month` = n, `Balance` = new_Balance)
  annuity_Table <- rbind(annuity_Table, new_row)
  
}

head_annuity_table <- head(annuity_Table)
tail_annuity_table <- tail(annuity_Table)
```

**Let's see a few records and the graphical representation.**

```{r, echo=FALSE, results='asis'}
print(xtable(head_annuity_table, caption='First records in the succession with no decimals', digits=0, floating=FALSE),hline.after=c(0,6), include.rownames=FALSE, type=type, comment=FALSE)
```

```{r, echo=FALSE, results='asis'}
print(xtable(tail_annuity_table, caption='Last records in the succession with no decimals', digits=0, floating=FALSE), hline.after=c(0,6), include.rownames=FALSE, type=type, comment=FALSE)
```

```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE, results='hide'}
p <- plot_ly(annuity_Table, 
             x = ~Month, 
             y = ~Balance,
             type = 'scatter', 
             mode = 'lines')

```

```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE, results='hide'}
pdf <- plot(annuity_Table$Month, 
            annuity_Table$Balance, 
            type="l", 
            col="blue",
            main='Annuity sequence',
            xlab="Month", 
            ylab="Balance")
pdf
```



### Answers

**Will the annuity run out of money? **

Yes, the annuity will run out of money.


**When?**

The annuity will run out of money in about 70 months from now on.

**What value will $a_n$ have when the annuity is depleted?**

$a_n$ will have a value of - \$338.09. That means that on their final withdrawal they will take home \$1000 - $338.09	= \$ `r 1000-338.09` since they gained some interest in the last month.





## Exercise \#9 Page Page 17.

The data in the accompanying table show the speed n (in increments of 5 mph) of an automobile and the associated distance $a_n$ in feet required to stop it once the brakes are applied. For instance, $n=6$ (representing $6 \times 5 = 30$ mph) requires a stopping distance of $a_6 = 47$ ft. 

```{r, echo=FALSE}
n <- c(1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16)
distance <- c(3,6,11,21,32,47,65,87,112,140,171,204,241,282,325,376)
stop_dist <- data.frame(`n` = n, `Distance` = distance)
```

```{r, echo=FALSE, results='asis'}
print(xtable(stop_dist, caption='n vs stopping distance', digits=0), hline.after=c(0,16), include.rownames=FALSE, type=type, comment=FALSE)
```

###  a.

**Calculate and plot the change $\Delta a_n$ versus n. Does the graph reasonably approximate a linear relationship?**

#### Solution

Now, the next table shows $\Delta a_n = a_{n+1} - a_n$.


```{r, echo=FALSE}
stop_dist <- data.table(stop_dist)
stop_dist[ , Delta := shift(Distance, type ="lead") - Distance]
```


```{r, echo=FALSE, results='asis'}
print(xtable(stop_dist, caption='Speed, stopping distance and delta', digits=0), hline.after=c(0,16), include.rownames=FALSE, type=type, comment=FALSE)
```


Let's plot our Delta vs our n values.


```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE, results='hide'}
pdf <- plot(stop_dist$n, 
            stop_dist$Delta, 
            type="p", 
            col="blue",
            main='Delta',
            xlab="n", 
            ylab="Delta in ft")
pdf
```


**Does the graph reasonably approximate a linear relationship?**

Yes, the graph reasonably approximate a linear relationship.


### b.

**Based on your conclusions in part (a), find a difference equation model for the stopping distance data. Test your model by plotting the errors in the predicted values against n. Discuss the appropriateness of the model.**

#### Solution

In order to solve this, we can basically calculate our stopping distance as follows:

$Distance_{n+1} = Distance_{n} + Delta_{n}$

Since we already know the values for the $Distance_n$; we could find a formula for Delta in terms of n since we concluded that is somehow linear.

That is:

$Delta = k \cdot n + b$

We could calculate k by finding the slope that passes in between the points $(n = 5, Delta = 15)$ and $(n = 6, Delta = 18)$.

$k \approx \frac{18-15}{6-5} \approx 3$

Now, if we want to substitute to find b, we could use as follows:

let's use the point $(n = 1, Delta = 3)$ and by replacing in the following formula:

$Delta = k \cdot n + b$

$3 = 3 \cdot 1 + b$

$0 = b$

Hence our final model formula for the Delta will be as follows:

$Delta_{n} = 3n$

Now that we have our $Delta_n$ in terms of n; we can proceed as follows:

$Predicted \: Distance_{n+1} = Predicted \: Distance_{n} + Delta_{n}$

Based on the above result; in order to obtain our final model formula, we could write as follows:

$Predicted \: Distance_{n+1} = Predicted \: Distance_{n} + 3n$

with $Predicted \: Distance_0 = 3$.

```{r, echo=FALSE}

for (i in stop_dist$n){
  if (i == 1){
    stop_dist$Predicted[1] <- stop_dist$Distance[1]
  }
  else {
    stop_dist$Predicted[i] <- stop_dist$Predicted[i-1] + 3*(i-1)
  }
}

stop_dist$Error <- stop_dist$Predicted - stop_dist$Distance
```


```{r, echo=FALSE, results='asis'}
print(xtable(stop_dist, caption='Speed, stopping distance, delta and predicted values', digits=0), hline.after=c(0,16), include.rownames=FALSE, type=type, comment=FALSE)
```

```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE, results='hide'}
plot(stop_dist$n, 
            stop_dist$Distance, 
            type="p", 
            col="blue",
            main='Predicted vs given value',
            xlab="n", 
            ylab="Stopping distance in ft")
points(stop_dist$n,stop_dist$Predicted,col="red")

```

The above plot, represents the real value in blue color and the predicted value in red.

```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE, results='hide'}
plot(stop_dist$n, 
            stop_dist$Error, 
            type="p", 
            col="blue",
            main='Error Predicted vs Real Value',
            xlab="n", 
            ylab="Error in ft")
```

The above graph represent the difference in between the actual reading and the predicted value employing the difference equation model for the stopping distance data.

In regards of the appropriateness of the model; it seems to be some how accurate with what seems to be a few outliers when the speed is high; this could be understandable and seems to predict the stopping distance fairly good; perhaps, this process could be improved using other methods.

## Exercise \#13 Page 34.

Consider the spreading of a rumor through a company of 1000 employees, all working in the same building. We assume that the spreading of a rumor is similar to the spreading of a contagious disease (see Example 3, Section 1.2) in that the number of people hearing the rumor each day is proportional to the product of the number who have heard the rumor previously and the number who have not heard the rumor. This is given by 

$r_{n+1} = r_n + k r_n(1000 - n)$

where k is a parameter that depends on how fast the rumor spreads and n is the number of days. Assume k = 0.001 and further assume that four people initially have heard the rumor. How soon will all 1000 employees have heard the rumor?

### Solution

From the above text we can collect the following information:

$r_0 = 4$

$k = 0.001$

By replacing in the above formula we obtain as follows:

$r_{n+1} = r_n + 0.001 r_n(1000 - n)$

$r_0 = 4$

Let's create a table of values based on the above dynamic system

```{r, echo=FALSE, }
n <- 0
rn <- 4
k <- 0.001

rn.df <- data.frame(n, rn)

while ((rn != 1000) & (rn != 999)){
  #rn <- trunc(rn + k * rn * (1000 - rn))
  #rn <- round(rn + k * rn * (1000 - rn),0)
  rn <- rn + k * rn * (1000 - rn)
  n <- n + 1
  rn.temp <- data.frame(n, rn)
  rn.df <- rbind(rn.df, rn.temp)
}
```

```{r, echo=FALSE}
rn.df
```


Let's plot our results:

```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE, results='hide'}
pdf <- plot(rn.df$n, 
            rn.df$rn, 
            type="p", 
            col="blue",
            main='Spread of rumor',
            xlab="n", 
            ylab="Number of Employees")
pdf
```

### Answer

**How soon will all 1000 employees have heard the rumor?**

If we take the decimals off from the results and we round down due to being person as a whole, our given function will never reach 1000 but 999 as the top value after 12 days.

If we round to the nearest decimals, then it will be reached after 11 days.

If we do not round any decimals, then it will take about 13 days.

So basically; it will be safe to say that the spread of the rumor will take from 11 to 13 days to spread to 1000 employees.

## Exercise \#6 Page Page 55.

An economist is interested in the variation of the price of a single product. It is observed that a high price for the product in the  market attracts more suppliers. However, increasing the quantity of the product supplied tends to drive the price down. Over time, there is an interaction between price and supply. The economist has proposed the following model, where $P_n$ represents the price of the product at year n, and $Q_n$ the quantity. Find the equilibrium values for this system.

$P_{n+1}=P_n-0.1(Q_n-500)$

$Q_{n+1}=Q_n+0.2(P_n-100)$

### Equilibrium values

If we call the equilibrium values (P, Q) then we must have $P=P_{n+1}=P_n$ and $Q=Q_{n+1}=Q_n$ simultaneously. Substituting into the above dynamical system yields:

$P=P-0.1(Q-500)$

$Q=Q+0.2(P-100)$

and by solving the above equations, we obtain

$Q=500$ and $P=100$

### a.

**Does the model make sense intuitively? What is the significance of the constants 100 and 500? Explain the significance of the sign of the constants -0.1 and 0.2.**

- Does the model make sense intuitively? **YES**.

- What is the significance of the constants 100 and 500? 

The constants 500  and  100  work  as  thresholds.

- Explain the significance of the sign of the constants -0.1 and 0.2.

Basically the constants -0.1 and 0.2 indicate as follows: The 0.2 indicate that the price has a positive change in the quantity if the price is over 100, while the -0.1 indicate that the quantity has a negative effect in the price if the quantity increases over 500.

### b.

**Test the initial condition in the following table and predict the long-term behaviour.**

```{r, echo=FALSE}
Case <- c('Case A','Case B','Case C','Case D')
Price <- c(100,200,100,100)
Quantity <- c(500,500,600,400)
t.df <- data.frame(Price,Quantity, row.names = Case)
```

```{r, echo=FALSE, results='asis'}
print(xtable(t.df, caption='', digits=0, floating=FALSE), hline.after=c(0,4), include.rownames=TRUE, type=type, comment=FALSE)
```

Please note that I have written a small script that will calculate our cases given our cases.

```{r, echo=FALSE}
runCase <- function(Case,P,Q,n_tests){
  n <- 0
  t.case <- data.frame(n,'Price' = P, 'Quantity' = Q)
  
  for (i in 0:(n_tests-2)){
    n <- i+1
    Pn1 <- t.case$Price[n] - 0.1 * (t.case$Quantity[n] - 500)
    Qn1 <- t.case$Quantity[n] + 0.2 * (t.case$Price[n] - 100)
    df <- data.frame(n,'Price' = Pn1, 'Quantity' = Qn1)
    t.case <- rbind(t.case, df)
  }
  plot(t.case$Quantity ~ t.case$n, 
       type = "p", 
       main = Case, 
       xlab="n", 
       ylab="Price & Quantity", 
       col = "blue", 
       pch=20, 
       xlim=c(-1,n_tests), 
       ylim=c(0,1000))
  lines(t.case$n, t.case$Price, col="red", type="p", pch=17)
  legend("topright", 
         c("Quantity","Price"),
         #pch=c(20,17),
         fill=c("blue","red")
         )
  
  tail(t.case, 10)
}
```

- **Case A**

With 10 iterations

```{r}
runCase("Case A",P = 100, Q = 500, 10)
```


With 100 iterations

```{r}
runCase("Case A",P = 100, Q = 500, 100)
```

With 1000 iterations

```{r}
runCase("Case A",P = 100, Q = 500, 1000)
```

- **Case B**

With 10 iterations

```{r}
runCase("Case B",P = 200, Q = 500, 10)
```

With 100 iterations

```{r}
runCase("Case B",P = 200, Q = 500, 100)
```

With 500 iterations

```{r}
runCase("Case B",P = 200, Q = 500, 500)
```

- **Case C**

With 10 iterations

```{r}
runCase("Case C",P = 100, Q = 600, 10)
```

With 100 iterations

```{r}
runCase("Case C",P = 100, Q = 600, 100)
```

With 500 iterations

```{r}
runCase("Case C",P = 100, Q = 600, 500)
```

- **Case D**

With 10 iterations

```{r}
runCase("Case D",P = 100, Q = 400, 10)
```

With 100 iterations

```{r}
runCase("Case D",P = 100, Q = 400, 100)
```

With 500 iterations

```{r}
runCase("Case D",P = 100, Q = 400, 500)
```

In conclusion, we can state that in this model, the equilibrium point is not stable.

# END.