---
title: "Project 1"
author: "Duubar Villalobos Jimenez"
date: "February 10, 2018"
output:
  pdf_document: default
  html_document: default
  prettydoc::html_pretty:
    highlight: github
    theme: leonids
    toc: yes
subtitle: CUNY MSDS DATA 608
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

#install.packages("install.load")
library(install.load)
install_load("plyr", "dplyr", "ggplot2", "scales", "ggthemes", "DT", "outliers")

```


**Principles of Data Visualization and Introduction to ggplot2**

I have provided you with data about the 5,000 fastest growing companies in the US, as compiled by Inc. magazine. lets read this in:

```{r}

inc <- read.csv("https://raw.githubusercontent.com/charleyferrari/CUNY_DATA_608/master/module1/Data/inc5000_data.csv", header= TRUE)

```

And lets preview this data:

```{r}

head(inc)

```

Summary:

```{r}

summary(inc)

```

Think a bit on what these summaries mean. Use the space below to add some more relevant non-visual exploratory information you think helps you understand this data:

```{r}

# Need to create individual summaries
Growth_Rate <- summary(inc$Growth_Rate)

Revenue <- summary(inc$Revenue)

Industry <- summary(inc$Industry)

Employees <- summary(inc$Employees)

City <- summary(inc$City)

State <- summary(inc$State)
```

## Question 1

Create a graph that shows the distribution of companies in the dataset by State (ie how many are in each state). There are a lot of States, so consider which axis you should use. This visualization is ultimately going to be consumed on a 'portrait' oriented screen (ie taller than wide), which should further guide your layout choices.

```{r, echo=TRUE, warning=FALSE, error=FALSE, cache=FALSE, results='hide', message=FALSE}
# Answer Question 1 here
my.data <- inc %>%
            group_by(State) %>%
            summarise('Count' = n()) %>%
            arrange(desc(`Count`))

# Basic plot
p <- ggplot(my.data, aes(x = reorder(State, Count), y = Count)) +
      geom_bar(stat='identity') +
      coord_flip() +
      xlab("State") + 
      ylab("Count") + 
      ggtitle("Counts by State") + 
      theme(plot.title = element_text(hjust = 0.5)) +
      scale_y_continuous() + 
      theme_bw()

```

```{r, echo=FALSE, warning=FALSE, error=FALSE, cache=FALSE, message=FALSE, fig.width=10, fig.height=11}
p
```

## Question 2

Lets dig in on the state with the 3rd most companies in the data set. Imagine you work for the state and are interested in how many people are employed by companies in different industries. Create a plot that shows the average and/or median employment by industry for companies in this state (only use cases with full data, use R's `complete.cases()` function.) In addition to this, your graph should show how variable the ranges are, and you should deal with outliers.


### State with the 3rd most companies

In order to get the State with the 3rd most companies, we can select as follows:

```{r}
# Answer Question 2 here
x = arrange(my.data,desc(Count))
x1 <- x[3:3,]  # Return the State with the 3rd most companies
```

From the above code we see that the State with the 3rd most companies is as follows:

```{r, echo=FALSE}
data.frame(x1)
```


```{r, echo=TRUE, warning=FALSE, error=FALSE, cache=FALSE, results='hide', message=FALSE}
# Subsetting data for the selected State
ind_by_state <- subset(inc, State == as.character(x1$State[1]))

# Complete Cases
ind_by_state <- ind_by_state %>%
                filter(complete.cases(Employees))

# Identifying outliers
ind_outliers <- ggplot(ind_by_state, aes(Industry, Employees)) +
                geom_boxplot() +
                coord_flip() +
                xlab("Industry") + 
                ylab("Number of Employees (With Outliers)") + 
                ggtitle("Number of Employees by Industry") + 
                theme(plot.title = element_text(hjust = 0.5)) +
                scale_y_continuous() + 
                theme_bw()
```

```{r, echo=FALSE, warning=FALSE, error=FALSE, cache=FALSE, message=FALSE, fig.width=10, fig.height=6}
ind_outliers 
```

Discarding Outliers (We will repeat the process 6 times in this case in order to identify if an outlier is detected and confirmed by statistical tests, this function can remove it or replace by sample mean or median. )

```{r}
# https://cran.r-project.org/web/packages/outliers/outliers.pdf

ind_no_outliers <- rm.outlier(ind_by_state$Employees, fill = TRUE, median = TRUE, opposite = FALSE)
ind_by_state$Emp_No_Outliers <- ind_no_outliers

ind_no_outliers <- rm.outlier(ind_by_state$Emp_No_Outliers, fill = TRUE, median = TRUE, opposite = FALSE)
ind_by_state$Emp_No_Outliers <- ind_no_outliers

ind_no_outliers <- rm.outlier(ind_by_state$Emp_No_Outliers, fill = TRUE, median = TRUE, opposite = FALSE)
ind_by_state$Emp_No_Outliers <- ind_no_outliers

ind_no_outliers <- rm.outlier(ind_by_state$Emp_No_Outliers, fill = TRUE, median = TRUE, opposite = FALSE)
ind_by_state$Emp_No_Outliers <- ind_no_outliers

ind_no_outliers <- rm.outlier(ind_by_state$Emp_No_Outliers, fill = TRUE, median = TRUE, opposite = FALSE)
ind_by_state$Emp_No_Outliers <- ind_no_outliers

ind_no_outliers <- rm.outlier(ind_by_state$Emp_No_Outliers, fill = TRUE, median = TRUE, opposite = FALSE)
ind_by_state$Emp_No_Outliers <- ind_no_outliers
```

```{r, echo=TRUE, warning=FALSE, error=FALSE, cache=FALSE, results='hide', message=FALSE}

# Identifying outliers
ind_outliers <- ggplot(ind_by_state, aes(Industry, Emp_No_Outliers)) +
                geom_boxplot() +
                coord_flip() +
                xlab("Industry") + 
                ylab("Number of Employees (No Outliers)") + 
                ggtitle("Number of Employees by Industry") + 
                theme(plot.title = element_text(hjust = 0.5)) +
                scale_y_continuous() + 
                theme_bw()
```

Result after 6 iterations; Outliers have been replaced by the Median value.

```{r, echo=FALSE, warning=FALSE, error=FALSE, cache=FALSE, message=FALSE, fig.width=10, fig.height=6}
ind_outliers 
```


Obtaining data with outliers

```{r, echo=TRUE, warning=FALSE, error=FALSE, cache=FALSE, results='hide', message=FALSE}
# Obatining  summary data for the state
my.data <- ind_by_state %>%
            group_by(Industry) %>%
            summarise('Count' = n(), 
                      'N_Employees' = sum(Employees), 
                      'Average' = round(mean(Employees),0), 
                      'Median' = round(median(Employees),0)) %>%
            arrange(desc(`Count`))

my.data <- data.frame(my.data)

# Basic plot Company Counts by Industry
p1 <- ggplot(my.data, aes(x = reorder(Industry, Count), y = Count)) +
      geom_bar(stat='identity') +
      coord_flip() +
      xlab("Industry") + 
      ylab("Count") + 
      ggtitle("Company Counts by Industry") + 
      theme(plot.title = element_text(hjust = 0.5)) +
      scale_y_continuous() + 
      theme_bw()
      
# Basic plot Number of Employees by Industry
p2 <- ggplot(my.data, aes(x = reorder(Industry, N_Employees), y = N_Employees)) +
      geom_bar(stat='identity') +
      coord_flip() +
      xlab("Industry") + 
      ylab("Count") + 
      ggtitle("Number of Employees by Industry") + 
      theme(plot.title = element_text(hjust = 0.5)) +
      scale_y_continuous() + 
      theme_bw()

# Basic plot Average Number of Employees by Industry
p3 <- ggplot(my.data, aes(x = reorder(Industry, Average), y = Average)) +
      geom_bar(stat='identity') +
      coord_flip() +
      xlab("Industry") + 
      ylab("Count") + 
      ggtitle("Average Number of Employees by Industry") + 
      theme(plot.title = element_text(hjust = 0.5)) +
      scale_y_continuous() + 
      theme_bw()
      
# Basic plot Median Number of Employees by Industry
p4 <- ggplot(my.data, aes(x = reorder(Industry, Median), y = Median)) +
      geom_bar(stat='identity') +
      coord_flip() +
      xlab("Industry") + 
      ylab("Count") + 
      ggtitle("Median Number of Employees by Industry") + 
      theme(plot.title = element_text(hjust = 0.5)) +
      scale_y_continuous() + 
      theme_bw()

```

Let's have a visual of the first few rows of the data:

```{r}
my.data
```

Graphical representation iof the data.

```{r, echo=FALSE, warning=FALSE, error=FALSE, cache=FALSE, message=FALSE, fig.width=10, fig.height=6}
p1
p2
p3
p4
```

Obtaining data with NO outliers

```{r, echo=TRUE, warning=FALSE, error=FALSE, cache=FALSE, results='hide', message=FALSE}
# Obatining  summary data for the state
my.data <- ind_by_state %>%
            group_by(Industry) %>%
            summarise('Count' = n(), 
                      'N_Employees' = sum(Emp_No_Outliers), 
                      'Average' = round(mean(Emp_No_Outliers),0), 
                      'Median' = round(median(Emp_No_Outliers),0)) %>%
            arrange(desc(`Count`))

my.data <- data.frame(my.data)

# Basic plot Company Counts by Industry
p1 <- ggplot(my.data, aes(x = reorder(Industry, Count), y = Count)) +
      geom_bar(stat='identity') +
      coord_flip() +
      xlab("Industry") + 
      ylab("Count") + 
      ggtitle("Company Counts by Industry") + 
      theme(plot.title = element_text(hjust = 0.5)) +
      scale_y_continuous() + 
      theme_bw()
      
# Basic plot Number of Employees by Industry
p2 <- ggplot(my.data, aes(x = reorder(Industry, N_Employees), y = N_Employees)) +
      geom_bar(stat='identity') +
      coord_flip() +
      xlab("Industry") + 
      ylab("Count") + 
      ggtitle("Number of Employees by Industry") + 
      theme(plot.title = element_text(hjust = 0.5)) +
      scale_y_continuous() + 
      theme_bw()

# Basic plot Average Number of Employees by Industry
p3 <- ggplot(my.data, aes(x = reorder(Industry, Average), y = Average)) +
      geom_bar(stat='identity') +
      coord_flip() +
      xlab("Industry") + 
      ylab("Count") + 
      ggtitle("Average Number of Employees by Industry") + 
      theme(plot.title = element_text(hjust = 0.5)) +
      scale_y_continuous() + 
      theme_bw()
      
# Basic plot Median Number of Employees by Industry
p4 <- ggplot(my.data, aes(x = reorder(Industry, Median), y = Median)) +
      geom_bar(stat='identity') +
      coord_flip() +
      xlab("Industry") + 
      ylab("Count") + 
      ggtitle("Median Number of Employees by Industry") + 
      theme(plot.title = element_text(hjust = 0.5)) +
      scale_y_continuous() + 
      theme_bw()

```

Let's have a visual of the first few rows of the data:

```{r}
my.data
```

Graphical representation iof the data (Since the Median and Mean are the same, I will present only one graphic).

```{r, echo=FALSE, warning=FALSE, error=FALSE, cache=FALSE, message=FALSE, fig.width=10, fig.height=6}
p1
p2
p3
p4
```

## Question 3

Now imagine you work for an investor and want to see which industries generate the most revenue per employee. Create a chart that makes this information clear. Once again, the distribution per industry should be shown.

```{r}

# Answer Question 3 here
inc$Rev_by_Emp <- inc$Revenue / inc$Employees

# Selecting the #3 State  (I am assuming, we are still working with the third state with most companies)
ind_by_state <- subset(inc, State == as.character(x1$State[1]))

# Complete Cases
ind_by_state <- ind_by_state %>%
                filter(complete.cases(Revenue, Employees))

# Obatining  summary data for the state
my.data <- ind_by_state %>%
            group_by(Industry) %>%
            summarise('Count' = n(), 
                      'N_Employees' = sum(Employees), 
                      'Tot_Revenue' = sum(Revenue)) %>%
            arrange(desc(`Tot_Revenue`))

my.data$Emp_Rev <- my.data$Tot_Revenue / my.data$N_Employees

my.data = arrange(my.data,desc(Emp_Rev))

my.data <- data.frame(my.data)
```

```{r}
my.data
```

```{r, echo=TRUE, warning=FALSE, error=FALSE, cache=FALSE, results='hide', message=FALSE}

# Basic plot Number of Revenue per Employee by Industry
p <- ggplot(my.data, aes(x = reorder(Industry, Emp_Rev), y = Emp_Rev)) +
      geom_bar(stat='identity') +
      coord_flip() +
      xlab("Industry") + 
      ylab("Revenue by Employee") + 
      ggtitle("Employee Revenue by Industry (New York)") + 
      theme(plot.title = element_text(hjust = 0.5)) +
      scale_y_continuous(labels = scales::dollar) +
      theme_bw()
```

```{r, echo=FALSE, warning=FALSE, error=FALSE, cache=FALSE, message=FALSE, fig.width=10, fig.height=6}
p
```


For the whole Country


```{r}

# Answer Question 3 here
inc$Rev_by_Emp <- inc$Revenue / inc$Employees

# Selecting the #3 State  (I am assuming, we are still working with all of the states)
ind_by_state <- inc

# Complete Cases
ind_by_state <- ind_by_state %>%
                filter(complete.cases(Revenue, Employees))

# Obatining  summary data for the state
my.data <- ind_by_state %>%
            group_by(Industry) %>%
            summarise('Count' = n(), 
                      'N_Employees' = sum(Employees), 
                      'Tot_Revenue' = sum(Revenue)) %>%
            arrange(desc(`Tot_Revenue`))

my.data$Emp_Rev <- my.data$Tot_Revenue / my.data$N_Employees


my.data <- data.frame(my.data)
```

```{r}
my.data
```

```{r, echo=TRUE, warning=FALSE, error=FALSE, cache=FALSE, results='hide', message=FALSE}

# Basic plot Number of Revenue per Employee by Industry
p <- ggplot(my.data, aes(x = reorder(Industry, Emp_Rev), y = Emp_Rev)) +
      geom_bar(stat='identity') +
      coord_flip() +
      xlab("Industry") + 
      ylab("Revenue by Employee") + 
      ggtitle("Employee Revenue by Industry (All States)") + 
      theme(plot.title = element_text(hjust = 0.5)) +
      scale_y_continuous(labels = scales::dollar) + 
      theme_bw()
```

```{r, echo=FALSE, warning=FALSE, error=FALSE, cache=FALSE, message=FALSE, fig.width=10, fig.height=6}
p
```
