---
title: "Homework 5.2"
author: "Duubar Villalobos Jimenez  mydvtech@gmail.com"
date: "November 4, 2018"

output:
 pdf_document:
   highlight: tango
   toc: true
   toc_depth: 4
   number_sections: true
   df_print: kable
   
 html_document: default
 
 prettydoc::html_pretty:
  highlight: github
  theme: leonids
  toc: yes
  
fontsize: 10pt
geometry: margin=1in
subtitle: CUNY MSDS DATA 609
---

```{r, echo=FALSE, warning=FALSE, error=FALSE, cache=FALSE, results='hide', message=FALSE}

knitr::opts_chunk$set(echo = TRUE, fig.pos = 'H')

```


```{r, echo=FALSE, warning=FALSE, error=FALSE, cache=FALSE, results='hide', message=FALSE}

library(knitr)
library(png) 

```

# Problems

The below problems are taken from the text book:

A First Course in Mathematical Modeling, 5th Edition. Frank R. Giordano, William P. Fox, Steven B. Horton. ISBN-13: 9781285050904.

## Exercise \#3 Page 469

**The following data were obtained for the growth of a sheep population introduced into a new environment on the island of Tasmania (adapted from J. Davidson, "On the Growth of the Sheep Population in Tasmania," *Trans. R. Soc. S. Asutralia* 62(1938): 342-346).**

```{r, echo = FALSE}
t <- c(1814,1824,1834,1844,1854,1864)
p <- c(125,275,830,1200,1750,1650)

df <- data.frame(t, p)
df <- data.frame(t(df))

names(df) <- NULL
rownames(df) <- c('t(year)','P(t)')
df
```

### Estimates

**Make an estimate of M by graphing $P(t)$.**

#### Solution

First, lets plot our data.

```{r, echo=FALSE}
df <- data.frame(t, p)
plot(df$p ~ df$t, 
   type ='o', 
   col = 'blue',
   xlab = 't(year)',
   ylab = 'P(t)')
```

From the graph, we can observe how there seems to be a trend and the maximum value $M$ seems no to exceed $M = 1800$.

### Plots

**Plot $ln[P/(M-P)]$ against $t$. If a logistic curve seems reasonable, estimate $rM$ and $t^*$.**

#### Solution

Let's see our plot:

```{r, echo=FALSE}
M <- 1800
df$ln <- log(df$p/(M - df$p))
```


```{r, echo=FALSE}
plot(df$ln ~ df$t, 
   type ='o', 
   col = 'blue',
   xlab = 't(year)',
   ylab = 'ln[P/(M-P)]')
abline(lm(df$ln ~ df$t), col = "red", lty = 3)
```

From the above plot, there series of values, seems not to be approximated by a linear model. Hence, a logistic curve seems to be appropriate.

Now, since

$$ln \frac{P}{M-P} = rMt + C$$

We could estimate the slope $rM$ as follows:


```{r, echo=FALSE}
linear_model <- lm(df$ln ~ df$t)
summary(linear_model)
```

From the above results, we have:

$\beta_0 = -218.31833$

$\beta_1 = 0.11891$

Resulting in:

$$ln \frac{P}{M-P} = \beta_0 + \beta_1t$$

Which means that $\beta_0 = C$ and $\beta_1 = rM$

$\therefore C = -218.31833$

$\therefore rM = 0.11891$

Now, in order to find $t^*$, we perform as follows:

$$t^* = -\frac{C}{rM}$$

$$t^* = -\frac{-218.31833}{0.11891}$$


```{r, echo=FALSE}
C <- linear_model$coefficients[[1]]
rM <- linear_model$coefficients[[2]]
t_star <- -C / rM

```

$t^*$ = `r t_star`.

An alternative, will be by employing the following formula:

$$t^* = t_0 - \frac{1}{rM} ln \frac{P_0}{M-P_0}$$

$$t^* = 1814 - \frac{1}{0.11891} ln \frac{125}{1800-125}$$

```{r, echo=FALSE}
t_star1 <- t[1] - 1/rM * log(p[1]/(M-p[1]))
```

$t^*$ = `r t_star1`.


```{r, echo=FALSE}
e <- -rM * (t - t_star)
df$lcurve <- M/(1+exp(e))
```

```{r, echo=FALSE}
plot(df$lcurve ~ df$t, 
   type ='o', 
   lty = 3,
   col = 'red',
   xlab = 't(year)',
   ylab = 'P(t)')
par(new=TRUE)
plot(df$p ~ df$t, 
   type ='p', 
   col = 'blue',
   xlab = 't(year)',
   ylab = 'P(t)')
```

## Exercise \#6 Page 478

**Suggest other phenomena for which the model described in the text might be used.**

### Solution

Let's see what is our Problem identification:

*How can the doses and the time between doses be adjusted to maintain a safe but effective concentration of the drug in the blood?*

### Suggested phenomena

In my case, I think that it could be used for colony of bacteria growing in an environment with limited resources; let's say with lack of food, or space constraints on the size of the colony. In this case, I believe that it is not reasonable to expect the colony to grow exponentially; making the colony unable to grow larger than some maximum population P.

